Clustering in a hyperbolic model of complex networks.

### 2020-03-10 concentration_argument.tex

We make two important observations. First, $\Xi(y,y^\prime)$ is an increasing function in both arguments, for $y, y^\prime < R$ and $y + y^\prime < R$. Second, for all $y + y^\prime < R$, $h_1(y) \le h_3(y^\prime)$ and $h_2(y) \le h_3(y)$, while $h_3(y), h_3(y^\prime) < 1$, so that
\begin{equation}\label{eq:derivative_hyp_ball_Xi_bounds}
	2 e^{-(R - y - y^\prime)}h_1(y) h_2(y) \le \Xi(y,y^\prime) \le 2 e^{-(R - y - y^\prime)}.
\end{equation}
In particular, since $R-y$ is an increrasing function of $n$ uniformly on $0 < y < (1-\varepsilon)R$, there exists a $0 < \delta < 1$ such that $1/2 \le \Xi(y,y^\prime) < 2$ for all $y + y^\prime < R$ and $(1-\delta)(R-y) < y^\prime < R$ and $n$ large enough.

@@

Recalling the definition of $I_2(y)$ we have
\begin{align*}
	&\hspace{-30pt}\left|\int_0^{(1-\delta)(R- y)} \frac{\Phi(y,y^\prime)}{2} e^{-\alpha y^\prime} \dd y^\prime 
		- I_2(y)\right|\\
	&\le \int_0^{(1-\delta)(R- y)} \left|\frac{\Phi(y,y^\prime)}{2}- \frac{\frac{1}{2}e^{R/2} \Xi(y,y^\prime)}
		{\sqrt{1 - \left(1 - \Xi(y,y^\prime)\right)^2}}(1+\varphi_n(y,y^\prime)\right|e^{-\alpha y^\prime} \dd y^\prime.
		\numberthis \label{eq:derivative_hyp_ball_final_error}
\end{align*}
We will oroceed to bound the term inside the integral. For this we first note that for $0 \le y^\prime \le (1-\delta)(R-y)$,
\[
	\varphi_n(y,y^\prime) \le \frac{e^{-\delta(R-y)}}{1 - e^{-\delta(R-y)}}.
\]
and recall that $\Xi(y,y^\prime) \le 2 e^{-(R-y-y^\prime)}$. Moreover, since $x/\sqrt{1-(1-x)^2} = x/\sqrt{2x-x^2}$ is an increasing function and $e^{-(R - y - y^\prime)} \le e^{-\delta(R- y)}$ for $0 < y^\prime < (1-\delta)(R-y)$,
\[
	\frac{\frac{1}{2}e^{R/2} \Xi(y,y^\prime)}{\sqrt{1 - \left(1 - \Xi(y,y^\prime)\right)^2}}
	\le e^{R/2} \frac{e^{-\delta(R-y)}}{\sqrt{2e^{-\delta(R-y)} - e^{-2\delta(R-y)}}}.
\]
Next, recall that $\Phi(y,y^\prime) = \frac{1}{2}e^{R/2}\arccos(1-\Xi(y,y^\prime))$. Then, since $\Xi(y,y^\prime) < 1$ for all $y^\prime < (1-\delta)(R-y)$, $y < R$ and $n$ large enough, we have (see Lemma~\ref{lem:arccos_approx}),
\[
	\left|\frac{1}{2}\Phi(y,y^\prime) - \frac{\frac{1}{2}e^{R/2} \Xi(y,y^\prime)}{\sqrt{1 - \left(1 - \Xi(y,y^\prime)\right)^2}}\right| \le \frac{1}{2}\Phi(y,y^\prime) \Xi(y,y^\prime).
\]
for all $y^\prime < (1-\delta)(R - y)$ and $y < R$. Togehter these facts imply that for $n$ large enough
\begin{align*}
	&\hspace{-30pt}\left|\frac{\Phi(y,y)}{2} - \frac{\frac{1}{2}e^{R/2} \Xi(y,y^\prime)}{\sqrt{1 - \left(1 - 	
		\Xi(y,y^\prime)\right)^2}}(1+\varphi_n(y,y^\prime))\right|\\
	&\le \frac{\Phi(y,y^\prime)\Xi(y,y^\prime)}{2} 
		+ \frac{\frac{1}{2}e^{R/2} \Xi(y,y^\prime) \varphi_n(y,y^\prime)}
		{\sqrt{1 - \left(1 - \Xi(y,y^\prime)\right)^2}} \\
	&\le e^{-\delta(R-y)} \Phi(y,y^\prime) + \frac{e^{-\delta(R-y)}}{1-e^{-\delta(R-y)}}
		\frac{e^{R/2} e^{-\delta(R-y)}}{\sqrt{2e^{-\delta(R-y)} - e^{-2\delta(R-y)}}}\\
	&\le e^{-\delta(R-y)} \Phi(y,y^\prime) + \frac{e^{\frac{R}{2}}e^{-\frac{3}{2}\delta(R-y)}}
		{\left(1 - e^{-\delta(R-y)}\right)^{3/2}}
\end{align*}

>>>>

We make two important observations. First, $\Xi(y,y^\prime)$ is an increasing function in both arguments, for $y, y^\prime < R$ and $y + y^\prime < R$. Second, for all $y + y^\prime < R$, $h_1(y) \le h_3(y^\prime)$ and $h_2(y) \le h_3(y)$, while $h_3(y), h_3(y^\prime) < 1$, so that
\begin{equation}\label{eq:derivative_hyp_ball_Xi_bounds}
	2 e^{-(R - y - y^\prime)}h_1(y) h_2(y) \le \Xi(y,y^\prime) \le 2 e^{-(R - y - y^\prime)}.
\end{equation}
In particular, since $R-y$ is an increasing function of $n$ uniformly on $0 < y < (1-\varepsilon)R$, there exists a $0 < \delta < 1$ such that $1/2 \le \Xi(y,y^\prime) < 2$ for all $y + y^\prime < R$ and $(1-\delta)(R-y) < y^\prime < R$ and $n$ large enough.

@@

Recalling the definition of $I_2(y)$ we have
\begin{align*}
	&\hspace{-30pt}\left|\int_0^{(1-\delta)(R- y)} \frac{\Phi(y,y^\prime)}{2} e^{-\alpha y^\prime} \dd y^\prime 
		- I_2(y)\right|\\
	&\le \int_0^{(1-\delta)(R- y)} \left|\frac{\Phi(y,y^\prime)}{2}- \frac{\frac{1}{2}e^{R/2} \Xi(y,y^\prime)}
		{\sqrt{1 - \left(1 - \Xi(y,y^\prime)\right)^2}}(1+\varphi_n(y,y^\prime)\right|e^{-\alpha y^\prime} \dd y^\prime.
		\numberthis \label{eq:derivative_hyp_ball_final_error}
\end{align*}
We will proceed to bound the term inside the integral. For this we first note that for $0 \le y^\prime \le (1-\delta)(R-y)$,
\[
	\varphi_n(y,y^\prime) \le \frac{e^{-\delta(R-y)}}{1 - e^{-\delta(R-y)}}.
\]
and recall that $\Xi(y,y^\prime) \le 2 e^{-(R-y-y^\prime)}$. Moreover, since $x/\sqrt{1-(1-x)^2} = x/\sqrt{2x-x^2}$ is an increasing function and $e^{-(R - y - y^\prime)} \le e^{-\delta(R- y)}$ for $0 < y^\prime < (1-\delta)(R-y)$,
\[
	\frac{\frac{1}{2}e^{R/2} \Xi(y,y^\prime)}{\sqrt{1 - \left(1 - \Xi(y,y^\prime)\right)^2}}
	\le e^{R/2} \frac{e^{-\delta(R-y)}}{\sqrt{2e^{-\delta(R-y)} - e^{-2\delta(R-y)}}}.
\]
Next, recall that $\Phi(y,y^\prime) = \frac{1}{2}e^{R/2}\arccos(1-\Xi(y,y^\prime))$. Then, since $\Xi(y,y^\prime) < 1$ for all $y^\prime < (1-\delta)(R-y)$, $y < R$ and $n$ large enough, we have (see Lemma~\ref{lem:arccos_approx}),
\[
	\left|\frac{1}{2}\Phi(y,y^\prime) - \frac{\frac{1}{2}e^{R/2} \Xi(y,y^\prime)}{\sqrt{1 - \left(1 - \Xi(y,y^\prime)\right)^2}}\right| \le \frac{1}{2}\Phi(y,y^\prime) \Xi(y,y^\prime).
\]
for all $y^\prime < (1-\delta)(R - y)$ and $y < R$. Togehter these facts imply that for $n$ large enough
\begin{align*}
	&\hspace{-30pt}\left|\frac{\Phi(y,y)}{2} - \frac{\frac{1}{2}e^{R/2} \Xi(y,y^\prime)}{\sqrt{1 - \left(1 - 	
		\Xi(y,y^\prime)\right)^2}}(1+\varphi_n(y,y^\prime))\right|\\
	&\le \frac{\Phi(y,y^\prime)\Xi(y,y^\prime)}{2} 
		+ \frac{\frac{1}{2}e^{R/2} \Xi(y,y^\prime) \varphi_n(y,y^\prime)}
		{\sqrt{1 - \left(1 - \Xi(y,y^\prime)\right)^2}} \\
	&\le e^{-\delta(R-y)} \Phi(y,y^\prime) + \frac{e^{-\delta(R-y)}}{1-e^{-\delta(R-y)}}
		\frac{e^{R/2} e^{-\delta(R-y)}}{\sqrt{2e^{-\delta(R-y)} - e^{-2\delta(R-y)}}}\\
	&\le e^{-\delta(R-y)} \Phi(y,y^\prime) + \frac{e^{\frac{R}{2}}e^{-\frac{3}{2}\delta(R-y)}}
		{\left(1 - e^{-\delta(R-y)}\right)^{3/2}}
\end{align*}
